---
title       : Earned Run Average (E.R.A.) Presentation
subtitle    : Developing Data Products Course Project
author      : Charles Westby
job         : Upcoming Data Analyst
framework   : io2012       # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, shiny, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## What is an Earned Run Average (E.R.A.)?

An E.R.A. is a statistic that is fundamental in the game of baseball. It is a calculation that tells how many Earned Runs a pitcher gives up on average per baseball game or 9 innings pitched.  
  
```{r opts.label = 'shiny', warning=FALSE, cache=FALSE, echo=FALSE}

library(slidifyLibraries)
library(shiny)
shinyUI(pageWithSidebar(

  titlePanel("Earned Run Average (E.R.A.)"),

  sidebarPanel(
    numericInput(inputId = "number1", label = "Earned Runs Allowed", value = 0, min = 0),
    numericInput(inputId = "number2", label = "Innings Pitched", value = 0, min = 0),
    submitButton("Calculate")
    ),

    mainPanel(
      p("Earned Runs Allowed"),
      textOutput("number1"),
      p("Innings Pitched"),
      textOutput("number2"),
      p("Earned Run Average (E.R.A)"),
      textOutput("text3")
    )
  )
)

shinyServer(function(input, output) {

  output$number1 <- renderText({input$number1})
  output$number2 <- renderText({input$number2})
  output$text3 <- renderText({round(input$number1/input$number2*9, digits = 2)})


  })



```
  




--- 


## What you need to calculate an Earned Run Average (E.R.A.)?

$$Earned.Run.Average=\frac{Earned.Runs}{Innings.Pitched}* 9$$  

* Earned Runs - An Earned Run is a Run that is given up when an Error in the field is not committed. If an Error in the field is committed, the Run is scored an Uneared Run. When in question, an Earned Run is decided by the Official Scorer of the game. 

* Innings Pitched - Innings pitched is how many in innings a pitcher has thrown up until that point or for the entire season. Each inning is measured by three outs. Therefore, if a pitcher has made it through 7 innings and has recorded one out, he has thrown 7.1 innings. If he has recorded two outs, he has thrown 7.2 innings.  

* The answer is then multiplied by 9 because there are 9 innings in a standard regulation baseball game. If the amount of innings per game is different in your league, then that number should be substituted for 9. 



---


## Calculate the Earned Run Average (E.R.A)

If you click the link, you will be sent to a page where you can calculate E.R.A.'s

http://cwestby.shinyapps.io/Shiny_App/  
  
R performs for following calculation 
``` {r eval= FALSE, tidy= TRUE} 
  output$text3 <- renderText({round(input$earned_runs/input$innings_pitched*9, digits = 2)})

  ```


### You know the Earned Run Average (E.R.A.). Now what? 
  
Once you calculate an E.R.A., there are several things you can do with the statistic. Along with other statistics, an E.R.A. can help compare pitchers to one another. Pitchers are often broken up into different classes. The most basis classes are Starting Pitchers and Relief Pitchers. Starting Pitchers are usually compared with other Starting Pitchers. The same goes for Relief Pitchers. This app helps facilitate the process.

---


## Why calculate an Earned Run Average (E.R.A.)?

* In the era of sabermetrics and baseball analytics, the E.R.A. is stil a valuable stat.  

* Much of the focus is now on pitches thrown per inning, which is good for determining a pitcher's efficiency. 

* Another critical stat that is new to the public is a pitcher's WHIP, or Walks and Hits per inning pitched. This stat shows how often a pitcher allows hitters on base. Pitchers with high WHIP's tend to have high E.R.A.'s.  

* However, an E.R.A. is still valuable because it shows how many runs a pitcher tends to give up when he pitches. In the end, wins and losses are measured in runs.





